---
title: "Homework 7"
subtitle: "Regression Analysis"
author: "Alexander Durbin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(gridExtra)
set.seed(12345)
```

# Linear Spline

- For *k* knots, $\lambda=0$ fits a straight line between each knot, so that the fitted spline better follows the curvature of the data. As $\lambda$ increases, the fitted spline becomes more segmented and does not follow the curvature of the data. Large $\lambda$ fits an almost straight line to the data.
- For fixed $\lambda$, increasing the number of knots the spline to fit more, smaller line segments that better fit the curvature of the data. 

```{r}

storage <- list("0" = NULL)

for(num_knots in c(5, 10, 20, 30)){
  for(l in c(0, 10, 15, 20, 50, 1000)){
    
    # generate data
    x <- runif(n = 1000, min = 0, max = 3*pi/2)
    e <- rnorm(n = 1000, mean = 0, sd = 0.1)
    y <- sin(x + 0.1) + e
    df <- data.frame(x, e, y)
    
    # get knot locations
    K <- num_knots
    k <- 1:K
    knots <- quantile(x, probs = (k+1)/(K+2))
    
    # make linear spline matrix
    X <- cbind(1, x)
    for(i in k){
      X <- cbind(X, (x - knots[i]) * as.numeric(x > knots[i]))
    }
    
    # make D
    D <- diag(c(0, 0, rep(1, K)))
    
    # set smoothing parameter
    lambda <- l
    
    # get beta hat
    beta_hat <- solve(t(X) %*% X + lambda * D, t(X) %*% y)
    
    # get estimates
    y_hat <- X %*% beta_hat
    df$y_hat <- y_hat
    
    # plot
    p <- ggplot(data = df) +
      geom_point(mapping = aes(x = x, y = y), alpha = 0.3, show.legend = FALSE) +
      geom_vline(xintercept = knots, alpha = 0.3) +
      stat_function(mapping = aes(color = "Truth"), 
                    fun = function(x) {sin(x + 0.1)}, lwd = 1) +
      geom_line(mapping = aes(x = x, y = y_hat, color = "Spline")) +
      scale_color_manual(name = "Fitted", values = c("red", "blue")) +
      ggtitle(paste0("Linear spline with ", K,
                     " knots \nand smoothing parameter ", lambda))
    
    storage[[as.character(num_knots)]][[as.character(l)]] <- p
    
  }
}

# arrange plots
for(i in c(5, 10, 20, 30)){
  do.call("grid.arrange", c(storage[[as.character(i)]], ncol = 2))
}

```

# Quadratic Spline

- Quadratic spline better fits the data here. For *k* knots, $\lambda=0$ provides a very wiggly fit, with each line segment being either a concave or convex fit. This results in a line that follows the noise of the data a little too well. Increasing $\lambda=10,20,50$ provides a more smooth fit that is less affected by the noise. But for $\lambda=1000$, we are provided with a single quadratic fit that does not reflect the overall curvature.
- High *k* and small $\lambda$ gives splines that follow the data too closely and are very wiggly. For moderate $\lambda$, the fit is almost identical and very smooth. High $\lambda$ and high *k* again provide a single quadratic fit.

```{r}

storage <- list("0" = NULL)

for(num_knots in c(5, 10, 20, 30)){
  for(l in c(0, 10, 15, 20, 50, 1000)){
    
    # generate data
    x <- runif(n = 1000, min = 0, max = 3*pi/2)
    e <- rnorm(n = 1000, mean = 0, sd = 0.1)
    y <- sin(x + 0.1) + e
    df <- data.frame(x, e, y)
    
    # get knot locations
    K <- num_knots
    k <- 1:K
    knots <- quantile(x, probs = (k+1)/(K+2))
    
    # make linear spline matrix
    X <- cbind(1, x)
    for(i in k){
      X <- cbind(X, (x - knots[i])^2 * as.numeric(x > knots[i]))
    }
    
    # make D
    D <- diag(c(0, 0, rep(1, K)))
    
    # set smoothing parameter
    lambda <- l
    
    # get beta hat
    beta_hat <- solve(t(X) %*% X + lambda * D, t(X) %*% y)
    
    # get estimates
    y_hat <- X %*% beta_hat
    df$y_hat <- y_hat
    
    # plot
    p <- ggplot(data = df) +
      geom_point(mapping = aes(x = x, y = y), alpha = 0.3, show.legend = FALSE) +
      geom_vline(xintercept = knots, alpha = 0.3) +
      stat_function(mapping = aes(color = "Truth"), 
                    fun = function(x) {sin(x + 0.1)}, lwd = 1) +
      geom_line(mapping = aes(x = x, y = y_hat, color = "Spline")) +
      scale_color_manual(name = "Fitted", values = c("red", "blue")) +
      ggtitle(paste0("Quadratic spline with ", K,
                     " knots \nand smoothing parameter ", lambda))
    
    storage[[as.character(num_knots)]][[as.character(l)]] <- p
    
  }
}

# arrange plots
for(i in c(5, 10, 20, 30)){
  do.call("grid.arrange", c(storage[[as.character(i)]], ncol = 2))
}

```
