---
title: "Homework 4"
subtitle: "STAT5514: Regression Analysis"
author: "Alexander Durbin"
date: "`r Sys.Date()`"
output: pdf_document
size: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)
```

```{r, include = FALSE}

library(tidyverse)
library(broom)
library(lmtest)

```

```{r}

raw <- read.table('data.txt', header = TRUE)
y <- raw$Lab
x <- cbind(rep(1, nrow(raw)), raw$Field)

tol <- 1E-5

beta <- coefficients(lm(y ~ raw$Field))

dif <- 1

while (dif > tol) {
  
  old <- beta
  r <- y - x %*% old
  theta <- coefficients(glm(r^2 ~ raw$Field, family = Gamma(link = 'log')))
  W = diag(as.vector(1 / exp( x %*% theta)))
  beta <- solve(t(x) %*% W %*% x) %*% t(x) %*% W %*% y
  dif <- sum((old - beta)^2)
  
}

```

```{r}

cat("The simple linear regression coefficients are:\nIntercept: ", coefficients(lm(Lab ~ Field, data = raw))[1],"\nSlope: ", coefficients(lm(Lab ~ Field, data = raw))[2])

cat("The dual regression coefficients are:\nIntercept: ", beta[1],"\nSlope: ", beta[2])

```

We see that the slope coefficients are not significantly different, but there is a difference in the intercept coefficients. The Bruece-Pagan test identifies non-constant variance in this data, as do the residual plots. Additionally, the data is non-normal as well. Dual regression may not be an appropriate choice here, however. There is little change in the fitted models.


```{r, echo = TRUE}

bptest(lm(raw$Lab ~ raw$Field))
shapiro.test(raw$Lab)

```

```{r}

mod <- lm(Lab ~ Field, data = raw)
df <- augment(mod)
ggplot(data = df) + 
  geom_jitter(mapping = aes(x = .fitted, y = .resid)) +
  ggtitle("Residuals Versus Fitted") +
  xlab("Fitted") +
  ylab("Residuals") +
  geom_abline(intercept = 0, slope = 0)

```

```{r}

mod <- lm(Lab ~ Field, data = raw)
df <- augment(mod)
ggplot(data = df) + 
  geom_jitter(mapping = aes(x = .fitted, y = abs(.resid))) +
  ggtitle("|Residuals| Versus Fitted") +
  xlab("Fitted") +
  ylab("|Residuals|") +
  geom_abline(intercept = 0, slope = 0)

```

```{r}

ggplot(data = raw) +
  stat_qq(mapping = aes(sample = Lab)) +
  ggtitle("Normal QQ Plot") +
  xlab("Theoretical Quantiles") +
  ylab("Sample Quantiles")

```

```{r}

ggplot(data = raw) +
  geom_jitter(mapping = aes(x = Field, y = Lab)) +
  geom_smooth(mapping = aes(x = Field, y = Lab, linetype = 'lm'), method = 'lm', se = FALSE, fullrange = TRUE) +
  geom_abline(mapping = aes(linetype = 'dual`', slope = beta[2], intercept = beta[1])) +
  scale_linetype_manual(values = c('solid', 'dashed')) +
  ggtitle("Regression Fits")

```
